Clean data

### 2023-08-03 analysis/02_clean-data.Rmd
Included tidy data, where each choice alternative has its own row for each respondent

@@

```{r save-clean-data}
write_rds(data, here("data", "derived_data", "data.rds"))
write_rds(dummy_X, here("data", "derived_data", "hmnl_dummy_design.rds"))
write_rds(dummy_interactions_X, here("data", "derived_data", "hmnl_dummy_interactions.rds"))

>>>>

# Tidy joined data

Creating a tidy version of the joined data -- the cleaned data and the dummy-coded design.

```{r tidy-joined-data}
# Create a data frame with one row per respondent choice and a list-column for the design.
tidy_dummy_X <- tibble(
  id = rep(1:dim(dummy_X)[1], dim(dummy_X)[2]) %>% sort(),
  choice = rep(1:dim(dummy_X)[2], dim(dummy_X)[1]),
  X = vector(mode = "list", length = dim(dummy_X)[1] * dim(dummy_X)[2])
)
ind <- 0
for (i in 1:dim(dummy_X)[1]) {
  for (j in 1:dim(dummy_X)[2]) {
    ind = ind + 1
    tidy_dummy_X$X[[ind]] = dummy_X[i,j,,]
  }
}

# test_id = 425; test_choice = 9
# test_01 <- tidy_dummy_X |> 
#   filter(id == test_id, choice == test_choice) |> 
#   select(X) |> unnest(cols = c(X)) |> as.matrix()
# sum(test_01 != dummy_X[test_id,test_choice,,])

# Unnest each design matrix X.
tidy_dummy_X <- tidy_dummy_X %>%
  unnest(X) %>% 
  mutate(alt = rep(1:dim(dummy_X)[3], dim(dummy_X)[1] * dim(dummy_X)[2])) %>%
  relocate(alt, .after = choice)

# Pivot the data by choice and create a unique id for each row and join.
tidy_data <- tidy_dummy_X %>%
  left_join(
    data %>%
      pivot_longer(cols = contains("Q4"), names_to = "choice", values_to = "Y") %>%
      # select(id, choice, Y) |>
      separate(choice, into = c("temp", "choice")) %>%
      mutate(choice = as.integer(choice)) %>%
      select(-temp),
    by = c("id", "choice")) %>%
  relocate(Y, .after = alt)

tidy_data
```

@@

```{r save-clean-data}
write_rds(data, here("data", "derived_data", "data.rds"))
write_rds(tidy_data, here("data", "derived_data", "tidy_data.rds"))
write_rds(dummy_X, here("data", "derived_data", "hmnl_dummy_design.rds"))
write_rds(dummy_interactions_X, here("data", "derived_data", "hmnl_dummy_interactions.rds"))
